---
title: "Milwaukee Crime"
author: "Chris DeAngelis, CFA"
date: "July 23, 2019"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

<style type="text/css">

h1.title {
  font-size: 38px;
  color: Black;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: Black;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: Black;
  text-align: center;
}
</style>

### Objective
  

First, I setup packages to run my analysis.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) # Tidying, manipulating, and charting data
library(lemon) # Added features to ggplot2
library(RColorBrewer) # Adding more color features to ggplot2
library(flextable) # Rich formatting for tables
library(scales) # Formatting numbers
library(leaflet) # Drawing maps
library(rgdal)
library(sp) # Converting coordinates from UTM format to 

# Defining a standard theme across all charts
chart_theme <-   theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(hjust = 0.5, color = "gray", size = 9),
    panel.background = element_blank()
)

# COOL: https://towardsdatascience.com/the-data-driven-guide-to-crime-in-milwaukee-b6a373f898e9
```

### Sourcing data

Exploring Milwaukee's crime data on data.milwaukee.gov (https://data.milwaukee.gov/dataset/wibr).
```{r source, echo = FALSE}
raw_data <- read_csv("https://data.milwaukee.gov/dataset/e5feaad3-ee73-418c-b65d-ef810c199390/resource/87843297-a6fa-46d4-ba5d-cb342fb2d3bb/download/wibr.csv")

data <- raw_data %>%
          select(2:6,9,11:length(raw_data))
```

### Crime by Type

Milwaukee's crime seems to have fallen consistently since 2005. This seems to be a broader trend across all crime types. 
```{r crime_type_setup}
# Identify instances of each crime type by year
crimes_count <- data %>%
                select(2, 11:length(data)) %>%
                #filter(ReportedYear == "2019") %>%
                group_by(ReportedYear) %>%
                summarize(Arson = sum(Arson), AssaultOffense = sum(AssaultOffense), Burglary = sum(Burglary), CriminalDamage = sum(CriminalDamage), Homicide = sum(Homicide), LockedVehicle = sum(LockedVehicle), Robbery = sum(Robbery), SexOffense = sum(SexOffense), Theft = sum(Theft), VehicleTheft = sum(VehicleTheft)) %>%
                gather(Crimes, Count, -ReportedYear)

# Defining chart colors
purples_colors <- brewer.pal(9, "Purples") 
purples_colors <- colorRampPalette(purples_colors)(length(unique(crimes_count$Crimes)))

# Look at overall trend in crime
ggplot(crimes_count, aes(x = ReportedYear, y = Count/1000, fill = Crimes)) +
  geom_bar(position = 'stack', show.legend = TRUE, stat = 'identity') +
  scale_fill_manual(values = purples_colors) +
  labs(title = "Milwaukee Crime Trending Downward", x = "", y = "Instances (000's)", subtitle = "Chris DeAngelis, CFA") +
  chart_theme

# Look at crime by type over year
ggplot(crimes_count, aes(x = ReportedYear, y = Count/1000)) +
  geom_bar(stat = 'identity', fill = purples_colors[length(purples_colors)]) +
  facet_rep_wrap(Crimes ~ ., scales = "fixed", repeat.tick.labels = TRUE) +
  labs(title = "Milwaukee Crime by Type Over Time", x = "", y = "Instances (000's)", subtitle = "Chris DeAngelis, CFA") +
  chart_theme
```

### 2018 Crime

Assault & Offense still stands out as a 1/3 of all reported crime in Milwaukee.
```{r crime_2018}
# Look at total crime in 2018
current_crime <- data %>%
                  filter(ReportedYear == 2018) %>%
                  select(-c("ReportedDateTime", "ReportedYear", "POLICE", "WARD"))
                  
# Calculate Crimes by Zip Code
crime_by_zip <- current_crime %>%
                  select(-c(1,2, 5,6)) %>%
                  group_by(ZIP) %>%
                  summarize(Arson = sum(Arson), AssaultOffense = sum(AssaultOffense), Burglary = sum(Burglary), CriminalDamage = sum(CriminalDamage), Homicide = sum(Homicide), LockedVehicle = sum(LockedVehicle), Robbery = sum(Robbery), SexOffense = sum(SexOffense), Theft = sum(Theft), VehicleTheft = sum(VehicleTheft)) %>%
                  gather(crimes, count, -ZIP) %>%
                  mutate(percent = count / sum(count) * 100) %>%
                  arrange(desc(percent))

#test <- test[!is.na(test$V1) || is.na(test$v2),]
#test <- test[!is.na(test$V2),]

# CONVERSION NOT WORKING YET
points <- SpatialPoints(test, proj4string=CRS("+proj=longlat"))
hmm <- spTransform(points, CRS("+proj=longlat +datum=WGS84"))

#https://rpubs.com/walkerke/rstudio_gis


# current_crime[!is.na(current_crime$RoughX) || !is.na(current_crime$RoughY),] %>%
#   leaflet() %>%
#   addTiles() %>%
#   addMarkers(~as.numeric('RoughY'), ~as.numeric('RoughX'))
#   addCircles(color = 'red')
  
 # fitBounds() fits the view into the rectangle [lng1, lat1] – [lng2, lat2]


```

### Crime by Location
```{r location}


```

### Weapon of Choice

Looks at the top 10 most frequent and least frequent weapons used between 2005 and 2019. I don't know why but I expected nunchucks. I did not expect purses.
```{r weapons}
# Build
weapons <- unlist(strsplit(data$WeaponUsed, "/|\\,"))
weapons_df <- as.data.frame(table(weapons))
#weapons <- unlist(strsplit(data$WeaponUsed, "/"))
#weapons <- append(weapons,unlist(strsplit(data$WeaponUsed, ",")))
weapons_tidy <- weapons_df %>%
                  filter(!weapons %in% c("UNKNOWN", "UNKNO", "NA", "OTHER", "NONE", "11", "12", "14", "16", "20")) %>%
                  group_by(weapons) %>%
                  summarise(instances = sum(Freq)) %>%
                  mutate(percent = instances / sum(instances) * 100) %>%
                  arrange(desc(instances)) %>%
                  mutate_at(vars(instances), comma, accuracy = 1) %>%
                  mutate_at(vars(percent), comma, accuracy = 1, suffix = "%")

# Top 10 weapons used between 2005 and 2019
colnames(weapons_tidy) <- c("Top 10 Weapons", "Instances", "% of Total")
flextable(head(weapons_tidy, 10)) %>%
  bg(part = "header", bg = "black") %>%
  color(part = "header", color = "white") %>%
  align(part = "all", align = "center") %>%
  align(j = 1, align = "left") %>%
  autofit()

# Bottom 10 weapons used between 2005 and 2019
colnames(weapons_tidy) <- c("Bottom 10 Weapons", "Instances", "% of Total")
flextable(tail(weapons_tidy, 10)) %>%
  bg(part = "header", bg = "black") %>%
  color(part = "header", color = "white") %>%
  align(part = "all", align = "center") %>%
  align(j = 1, align = "left") %>%
  autofit()

```

